\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
First (rough) draft of section 5 done.

\section{Formulating Security with a Fully Abstract Overlay Semantics}
\label{sec:form-secur-with}
A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
Our proof uses a novel approach we call fully abstract overlay semantics.
The idea is to define a second operational semantics for programs in our target language.
This second semantics uses a different abstract machine and different run-time values, but it executes in lock-step with the original semantics and there is a very close correspondence between the state of both machines.

The main difference between the two semantics, is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.

The fact that \stktokens{} enforces LSE and WBCF is formulated as the theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
The theorem states that this function consitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than the corresponding components in the well-behaved overlay semantics.
In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics iff they are indistinguishable to arbitrary other components in the overlay semantics.

Our formal results are complicated by the fact that they only hold for well-behaved components that respect the basic rules of the calling convention, and also on a sane initial configuration of the system.
For example, the system should be setup such that seals that components use for constructing return pointers should not be shared with other components.
We envision distributing seals as a job for the linker, so this means our results depend on a linker to do this properly.
As another example, a seal that is used to construct a return pointer, can be used again to construct other return pointers for the same return point, but a different seal must be used for others.
Such seals should also never be passed to other components.
Such requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
The general pattern is that $\stktokens{}$ only protects components that do not shoot themselves in the foot by violating a few basic rules.
In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components to do certain unsafe things.

@@

% Back to high-level view
As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
However, there is no such assumption on the untrusted other components they interact with.
To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
Only instructions at these addresses are allowed to perform calls that push stack frames to the call stack, i.e.\ calls for which \srccm{} provides native LSE and WBCF guarantees.
The constant $\ta$ is passed around as a parameter of the \trgcm{} step relation.
Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.

Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ({\footnotesize $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is $26$, i.e.\ the number of instructions).
Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.

@@

Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
%The remaining defintions
\begin{definition}[Plugging a component into a context]
  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
\end{definition}
\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
  \label{def:contextual-equivalence}
  ~
  \begin{description}
  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
  \begin{equation*}
    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
  \end{equation*}
\end{description}
where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
\end{definition}
With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
% Full abstraction thm
\todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
\begin{theorem}
  \label{thm:full-abstraction}
  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
  \begin{gather*}
    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2}
  \end{gather*}
\end{theorem}
Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
In order to enforce these properties, components on \trgcm{} use \stktokens{}.
Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
\todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}

@@

\section{Proving full abstraction}
\label{sec:fa-proof}
\todo[inline]{Write small introduction. Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
% \begin{itemize}
% \item Logical relation
% \item FTLR
% \item Sketch high-level structure of the proof
% \end{itemize}
\todo[inline]{Describe figure here or save it for proof sketch subsection?}
\begin{figure}
  \centering
  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
    % \draw[help lines,yellow] (0,0) grid (10,7);
    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };

    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
  \end{tikzpicture}
  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
\end{figure}

\subsection{Logical relation}
\label{subsec:worlds}
The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
\todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
Unfortunately due to space considerations, we cannot give a comprehensive description of it.
Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensible description of a logical relation for a capability machine.
Finally, for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.

\subsubsection{Worlds}
Generally speaking, worlds are used in a logical relation to model the memory of a language.
\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds should model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to some regions that depend on the kind of sub-world.
where the regions differ for each of the sub-worlds.
Generally, the regions consist of a world indexed relation that specifies the memories allowed by the region and a region type.
The region type indicates how the region can be addressed which is important for linearity.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
The $\spatial$ region may seem odd as it not addressable, but all the relations in the logical relation are world indexed.
If the register-file relation and the memory relation both have worlds with the same ownership, then they can both contain linear capabilities that address the same memory.
If a spatial region is owned by the register-file relation, then it still has to be modelled in the memory which it can use a spatial region for.
To capture this, we define a relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint. We also define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
In a sense, linear capabilities are revocable, so the sub-worlds that allow spatial regions must also have a $\revoked$ region to model revocation.
Finally, the $\pure$ region also contains a seal interpretation function.
The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
We split the regions in two categories ($\Wor$ given below):
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
\end{multline*}
\begin{multline*}
  \Regions = 
    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup \\
    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
    \{\revoked\}
\end{multline*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the memory can contain every kind of region because it is addressable by both normal and linear capabilities.
The call stack and free stack should only be addressable by linear capabilities, so they only allow regions from $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.

@@

\subsection{Component logical relation}
The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
The component relation relates a component $(\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss)$ to itself when two conditions are satisfied.
First, when words that relate to them selves in the untrusted value relation are used to satisfy the imports, i.e.\ the words are placed on the import addresses in $\msdata$, and this data memory is combined with the code memory $\mscode$, then it forms a safe heap, i.e.\ it is in the $\mathcal{H}$ relation.
Second, the exports should always be safe to use which means that they must be in the untrusted value relation in any future world.

% \begin{multline*}
%   \lrcomp(W) =\\
%   \left\{\begin{aligned}
%       &\npair{\var{comp},\var{comp}} \;\mid \;\\
%       &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
%       &\qquad\text{For all } W' \future W \ldotp \\
%       &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
%       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
%       &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
%       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
%     \end{aligned}
%   \right\}\\
% \cup \left\{
%     \begin{multlined}
%       \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
%       \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
%       \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
%     \end{multlined}
%   \right\} 
% \end{multline*}

\begin{lemma}[FTLR for components]
  \label{lem:ftlr-comps}
  If
  \begin{itemize}
  \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
  \item $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$
  \item One of the following holds:
    \begin{itemize}
    \item $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component (see Section~\ref{sec:reasonability})
    \item $\dom(\comp.\mscode) \mathrel{\#} \ta$
    \end{itemize}
  \end{itemize}
  Then there exists a $W$ such that
  \begin{itemize}
  \item $\npair{(\comp,\comp)} \in \lrcomp(W)$
  \item $\dom(\pwheap)$ can be chosen to not include any finite set of region names 
  \item $\dom(\pwfree) =
    \dom(\pwpriv)=\emptyset$
  \end{itemize}
\end{lemma}
\begin{lemma}[Compatibility lemma for context plugging]
  \label{lem:compat-context-plug}
  If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
  $\npair[n']{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
\end{lemma}

\subsection{Executable configuration logical relation}

\begin{multline*}
  \lrec[\square,\gc = (\ta,\stkb)](W) = \\
\left\{
  \begin{array}{l}
     \npair{\left(
    (\ms_S,\reg_S,\stk,\ms_\stk),
    (\ms_T,\reg_T)\right)} \mid \\
    \quad \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
    \qquad \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
    \qquad \reg_S(\pcreg) \neq \retptrc(\_) \tand
    \reg_S(\rdata) \neq \retptrd(\_) \tand \\
    \qquad \nonExec{\reg_S(\rdata)} \tand  
     \nonExec{\reg_T(\rdata)} \tand \\
    \qquad \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
  \end{array}
\right\}
\end{multline*}
\begin{lemma}[Adequacy of execution configuration LR]
  \label{lem:adequacy}
  If $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$ and
 $i \leq n$ and
and $\Phi_S \sterm[i]{\gc}$,
  then $\Phi_T\term$.

  Also, if
 $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$ and
 $i \leq n$
and $\Phi_T \term[i]$,
  then $\Phi_S\sterm{\gc}$.
\end{lemma}

\subsection{Proof sketch}
\label{subsec:proof-sketch}
\begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
  \todo[inline]{Rewrite}
  % \item Consider first the upward arrow.
  %   Assume $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$.

@@

  First consider the right arrow:

>>>>

\section{Formulating Security with a Fully Abstract Overlay Semantics}
\label{sec:form-secur-with}
A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
Our proof uses a novel approach we call fully abstract overlay semantics.
In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
An overlay semantics is basically a different view on an existing semantics.
For instance, it may have new syntactic constructs and have special semantics for these things, but they correspond to something on the underlying machine.
In the overlay semantics we are about to define for \trgcm{}, an explicit stack pointer construct is part of the new view.
The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
When we have defined the overlay semantics, we prove full-abstraction for a compilation from the overlay semantics to the underlying semantics.
Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.
In our case, the compilation even turns out to be the identity function!

% protection against others purpose of cc
We can not show that \stktokens{} guarantee much if the security mechanisms of the capability machine are undermined by a poorly configured machine, e.g.\ the same seal is present in different components or the stack capability have aliases.
It is also possible to for a caller to undermine the security mechanisms during execution.
For example, the caller could give the return seal for a call to the callee.
\stktokens{} do not protect a component against itself as the component is assumed to behave reasonably.
In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components from unsafe behavior.

@@

% Back to high-level view
In \stktokens{} the base address of the memory $\stkb$ used for the stack is agreed upon upfront.
As \srccm{}'s view builds-in the stack, it needs to know the base address during execution.
From the perspective of a well-behaved component, other components can do whatever they want for their calls and local state as long as they cannot break WBCF and LSE for the well-behaved component.
Due to this, \srccm{} has a set of trusted addresses $\ta$ that are allowed to perform machine calls that push stack frames to the call stack.
To make these constants available during execution, they are passed around in the \trgcm{} step relation.
Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ({\footnotesize $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is $26$, i.e.\ the number of instructions).
Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.

@@

Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
%The remaining defintions
\begin{definition}[Plugging a component into a context]
  \label{def:comp-context-plugging}
  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
\end{definition}
\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
  \label{def:contextual-equivalence}
  ~
  \begin{description}
  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
  \begin{equation*}
    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
  \end{equation*}
\end{description}
where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
\end{definition}
With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
% Full abstraction thm
\todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
\begin{theorem}
  \label{thm:full-abstraction}
  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
  \begin{gather*}
    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2}
  \end{gather*}
\end{theorem}
Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
In order to enforce these properties, components on \trgcm{} use \stktokens{}.
Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
\todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}

@@

\section{Proving full abstraction}
\label{sec:fa-proof}
\todo[inline]{Write small introduction. Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
% \begin{itemize}
% \item Logical relation
% \item FTLR
% \item Sketch high-level structure of the proof
% \end{itemize}

\subsection{Logical relation}
\label{subsec:logical-relation}
The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
\todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
Unfortunately due to space considerations, we cannot give a comprehensive description of it.
Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensible description of a logical relation for a capability machine.
Finally, for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, worlds are used in a logical relation to model the memory of a language.
\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds should model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to some regions that depend on the kind of sub-world.
where the regions differ for each of the sub-worlds.
Generally, the regions consist of a world indexed relation that specifies the memories allowed by the region and a region type.
The region type indicates how the region can be addressed which is important for linearity.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
The $\spatial$ region may seem odd as it not addressable, but all the relations in the logical relation are world indexed.
If the register-file relation and the memory relation both have worlds with the same ownership, then they can both contain linear capabilities that address the same memory.
If a spatial region is owned by the register-file relation, then it still has to be modelled in the memory which it can use a spatial region for.
To capture this, we define a relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint. We also define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
In a sense, linear capabilities are revocable, so the sub-worlds that allow spatial regions must also have a $\revoked$ region to model revocation.
Finally, the $\pure$ region also contains a seal interpretation function.
The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
We split the regions in two categories ($\Wor$ given below):
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
\end{multline*}
\begin{multline*}
  \Regions = 
    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup \\
    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
    \{\revoked\}
\end{multline*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the memory can contain every kind of region because it is addressable by both normal and linear capabilities.
The call stack and free stack should only be addressable by linear capabilities, so they only allow regions from $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.

@@

\subsection{Full abstraction proof sketch}
\label{subsec:proof-sketch}
The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward.
The component relation relates components to them self which makes sense as well-formed components should contain no \srccm{} specific constructs.
We prove a number of lemmas about the relations.
The first is a FTLR for components that says (1) all well-formed untrusted components are related to themselves in the component relation, and (2)  all well-formed and reasonable trusted components are related to themselves in the component relation.
% \subsection{Component logical relation}
% The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
% The component relation relates a component $(\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss)$ to itself when two conditions are satisfied.
% First, when words that relate to them selves in the untrusted value relation are used to satisfy the imports, i.e.\ the words are placed on the import addresses in $\msdata$, and this data memory is combined with the code memory $\mscode$, then it forms a safe heap, i.e.\ it is in the $\mathcal{H}$ relation.
% Second, the exports should always be safe to use which means that they must be in the untrusted value relation in any future world.
%   \lrcomp(W) =\\
%   \left\{\begin{aligned}
%       &\npair{\var{comp},\var{comp}} \;\mid \;\\
%       &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
%       &\qquad\text{For all } W' \future W \ldotp \\
%       &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
%       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
%       &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
%       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
%     \end{aligned}
%   \right\}\\
% \cup \left\{
%     \begin{multlined}
%       \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
%       \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
%       \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
%     \end{multlined}
\begin{lemma}[FTLR for components]
  \label{lem:ftlr-comps}
  If
  \begin{itemize}
  \item $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$
  \item One of the following holds:
    \begin{itemize}
    \item $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component
    \item $\dom(\comp.\mscode) \mathrel{\#} \ta$
    \end{itemize}
  \end{itemize}
  Then there exists a $W$ such that
  $\npair{(\comp,\comp)} \in \lrcomp(W)$
\end{lemma}
The next is a compatibility lemma for plugging that says that given a context and component each related to itself in the component relation, then plugging the context with the component will give an executable configuration in the executable configuration relation.
\begin{lemma}[Compatibility lemma for context plugging]
  \label{lem:compat-context-plug}
  If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
  $\npair[n']{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
\end{lemma}
Finally, we have an adequacy lemma for the execution configuration relation.
This lemma says that if an \srccm{} configuration approximates a \trgcm{} configuration, and the \trgcm{} configuration terminates, then so does the \srccm{} configuration.
The symmetric case is also true.
%   \lrec[\square,\gc = (\ta,\stkb)](W) = \\
%      \npair{\left(
%     (\ms_S,\reg_S,\stk,\ms_\stk),
%     (\ms_T,\reg_T)\right)} \mid \\
%     \quad \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
%     \qquad \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
%     \qquad \reg_S(\pcreg) \neq \retptrc(\_) \tand
%     \reg_S(\rdata) \neq \retptrd(\_) \tand \\
%     \qquad \nonExec{\reg_S(\rdata)} \tand  
%      \nonExec{\reg_T(\rdata)} \tand \\
%     \qquad \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
\begin{lemma}[Adequacy of execution configuration logical relation]
  \label{lem:adequacy}
  If $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$ and
 $i \leq n$ and
and $\Phi_S \sterm[i]{\gc}$,
  then $\Phi_T\term$.

  Also, if
 $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$ and
 $i \leq n$
and $\Phi_T \term[i]$,
  then $\Phi_S\sterm{\gc}$.
\end{lemma}
We use these three lemmas to prove Theorem~\ref{thm:full-abstraction}
\begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
  The proof of both directions are very similar, so we only show the right direction.
  W.l.o.g.\ it suffices to show one direction of the contextual equivalence, so we assume ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
  The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.
  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that $\comp_1$ and $\comp_2$ are well-formed and reasonable.
  Further, by $\tconeq$ the context $\context$ must be well-formed.
  We prove arrow (1) in the figure by using the mentioned assumption about $\comp_1$ and $\context$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$, so by the assumption that $\comp_1$ and $\comp_2$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ which proves arrow (2) in the figure.
  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use that $\comp_2$ is well-formed and reasonable and that $\context$ is well-formed.
\end{proof}

\begin{figure}
  \centering
  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
    % \draw[help lines,yellow] (0,0) grid (10,7);
    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };

    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
  \end{tikzpicture}
  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
  \label{fig:fa-proof-sketch}
\end{figure}

% \subsection{Proof sketch}
% \label{subsec:proof-sketch}
% \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
  % \item Consider first the upward arrow.
  %   Assume $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$.

@@

%   First consider the right arrow:
